Data Analysis

### 2023-03-24 ms1_data_analysis.Rmd
Before introducing FUNGICIDE treatment analyses.

#### Complex model model
$\Delta$Biomass ~ 
SLA $\times$ site + SLA^2 $\times$ site + 
water.content $\times$ site + water.content^2 $\times$ site + 
perc.herb $\times$ site  + perc.herb^2 $\times$ site + 
richness + abundance + $\Delta$SLA + 
(Phylogenetic structure).

#### 1.1 Insecticide

```{r}
# Create data-set for a given treatment
deltaData <- data.frame()

@@

# The Focused model
# Only the predictive power of SLA, water content and percentage herbivory on deltaBio

@@

dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d,
                     data = wdata,
                     method = "ML",
                     correlation = corMartins(2.718282, 
                                              phy = myTree.rand.poly, 
                                              form = ~unified_names,
                                              fixed = T),
                     na.action = na.omit)

@@

dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d,
                     data = ndata,
                     method = "ML",
                     correlation = corMartins(2.718282, 
                                              phy = myTree.rand.poly, 
                                              form = ~unified_names,
                                              fixed = T),
                     na.action = na.omit)

@@

# The model
dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d,
                     data = ydata,
                     method = "ML",
                     correlation = corMartins(2.718282, 
                                              phy = myTree.rand.poly, 
                                              form = ~unified_names,
                                              fixed = T),
                     na.action = na.omit)

@@

# print the MSE
mse_focused_df <- rbind(mse_focused_df,  
                        data.frame(mse = with(ydata, mean((lrrtb - preds))^2),
                                   treatment = "Insecticide", site = "Yawan"))
```

```{r, results='asis'}
library(sjPlot)

@@

```{r, fig.cap = "Empirical and predicted values of a simple model of biomass change vs logarithm of SLA for each studied sites."}

@@

ggpubr::ggarrange(wpredPlot, npredPlot, ypredPlot,
                  nrow = 1)

```{r}
# Plot the relationship and add predicted values
ggplot(rbind(ydata, ndata, wdata), aes(y = lrrtb, x = lsla.d, color = site))+
  geom_point() + 
  geom_line(data = rbind(ydata, ndata, wdata), 
            aes(x = lsla.d, y = preds), lwd = 1)+
  geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15)

@@

>>>>

Explore that visually

#### Insecticide

##### Change in Biomass: log-respose ratio of a species' biomass

# Get the delta data ready for th e given subset vars (comparison)

# Create data-set for a given treatment
deltaData <- data.frame()

@@

Below I explore nonlinear relationships of the biomass log-ratios with species traits. Nonlinear relationships are introduced to allow for optimum trait values.

# Log-response ratios of total biomass vs of sla

# DeltaBiomass vs SLA at each site
slaPlot <- ggplot(deltaData, aes(x = lsla.d, y = lrrtb, color = site)) +
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ x + I(x^2))

# There might be some effect of site.
# DeltaBiomass vs SLA at each site
wcPlot <- ggplot(deltaData, aes(x = water.cont, y = lrrtb, color = site)) +
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ x + I(x^2))
# Seems like the effect is not very strong either

herbPlot <- ggplot(deltaData, aes(x = perc.herb, y = lrrtb, color = site)) +
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ x + I(x^2)) + 
  scale_x_continuous(trans = 'log2')

ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)

```
It seems that in some cases the nonlinear relationship may be important. Generally water content looks linear for each elevation. We will include this and build also a model that models log-ratio biomass vs water content as a linear relationship. We arrive to the final definition of the full model.

$\Delta$Biomass ~ SLA $\times$ site + SLA^2 $\times$ site + 
water.content $\times$ site + water.content^2 $\times$ site + 
perc.herb $\times$ site  + perc.herb^2 $\times$ site + 
richness + abundance + $\Delta$SLA + 
(Phylogenetic structure).

```{r}
simple.sla <- gls(
  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

just.sla <- gls(
  lrrtb ~ lsla.d*site + I(lsla.d^2)*site,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

AIC(just.sla, simple.sla)
anova(just.sla, simple.sla) # simple SLA is much bette
tab_model(simple.sla) # But only the lrr SLA is significant.

# Run the full model.
full.sla <- gls(
  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + water.cont*site + I(water.cont^2)*site + perc.herb*site + I(perc.herb^2)*site + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

# Assuming linear relationships
full.sla.linear <- gls(
  lrrtb ~ lsla.d*site + water.cont*site + perc.herb*site + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

anova(simple.sla, full.sla.linear) # full SLA linear is much better that the one using SLA
tab_model(full.sla.linear)

full.sla.linear.noint <- gls(
  lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla + site,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

anova(full.sla.linear, full.sla.linear.noint)
# Removing interactions does not improve the model significantly
tab_model(full.sla.linear.noint)

# Add interaction only to the water content and sla
full.sla.linear.intslawc <- gls(
  lrrtb ~ lsla.d*site + water.cont*site + perc.herb + richness + labu + lrrsla + site,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

anova(full.sla.linear.noint, full.sla.linear.intslawc)
# No effect
tab_model(full.sla.linear.noint)

full.sla.linear.intsla <- gls(
  lrrtb ~ lsla.d*site + water.cont + perc.herb + richness + labu + lrrsla + site,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

anova(full.sla.linear.noint, full.sla.linear.intsla)
# Any for of iteraction is not improving the model
tab_model(full.sla.linear.noint) # Thus the simplest model seems to be best.

# I further compare to the full.sla.linear.noint model.
# Only water content is modeled as linear
full.sla.custom <-  gls(
  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + water.cont*site + perc.herb*site + I(perc.herb^2)*site + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

anova(full.sla.linear.noint, full.sla.custom)

full.sla.custom.noint <-  gls(
  lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + perc.herb + I(perc.herb^2) + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

full.sla.linear.noint.nosite <- gls(
  lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

# Withouth a site it is even better
anova(full.sla.linear.noint, full.sla.linear.noint.nosite)

# The linear model remains the best one.
plot(full.sla.linear.noint.nosite) # This plot looks best
```

Table. Results of the selected model for the deltaBio.

```{r}
sjPlot::tab_model(full.sla.linear.noint.nosite)
```

Nevertheless it is not a strong correlation and does not explain a lot of variation in the biomass change.

```{r, fig.cap="Observed vs predicted values for the lsls.d and "}
preddata <- data.frame(lsla.d = seq(from = min(deltaData$lsla.d, na.rm = T), 
                                  to = max(deltaData$lsla.d, na.rm = T), 
                                  length = 100),
                       water.cont = mean(deltaData$water.cont, na.rm = T), 
                       perc.herb= mean(deltaData$perc.herb, na.rm = T), 
                       richness= mean(deltaData$richness, na.rm = T), 
                       labu= mean(deltaData$labu, na.rm = T),
                       lrrsla= mean(deltaData$lrrsla, na.rm = T))

preddatalrrsla <- data.frame(lsla.d = mean(deltaData$lsla.d, na.rm = T),
                       water.cont = mean(deltaData$water.cont, na.rm = T), 
                       perc.herb= mean(deltaData$perc.herb, na.rm = T), 
                       richness= mean(deltaData$richness, na.rm = T), 
                       labu= mean(deltaData$labu, na.rm = T),
                       lrrsla=  seq(from = min(deltaData$lrrsla, na.rm = T), 
                                  to = max(deltaData$lrrsla, na.rm = T), 
                                  length = 100))

visData <- deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label)

predvals <- AICcmodavg::predictSE.gls(full.sla.linear.noint.nosite, newdata = preddata,
                    newdtype = 'response')

preddata$predvals <- predvals$fit
preddata$predse <- predvals$se.fit

predvalslrrsla <- AICcmodavg::predictSE.gls(full.sla.linear.noint.nosite, newdata = preddatalrrsla,
                    newdtype = 'response')

preddatalrrsla$predvals <- predvalslrrsla$fit
preddatalrrsla$predse <- predvalslrrsla$se.fit

# For lsla.d
p1 <- ggplot(preddata, aes(x = lsla.d, y = predvals)) + 
  geom_line()+
  geom_ribbon(aes(ymin = predvals - predse,
                  ymax = predvals + predse), 
              alpha = 0.1) + 
  geom_point(data = visData, aes(x = lsla.d, y = lrrtb)) + 
  ggtitle("Log(SLA) vs Log-response ratio of a species' biomass")

# For lrrsla
p2 <- ggplot(preddatalrrsla, aes(x = lrrsla, y = predvals)) + 
  geom_line()+
  geom_ribbon(aes(ymin = predvals - predse,
                  ymax = predvals + predse), 
              alpha = 0.1) + 
  geom_point(data = visData, aes(x = lrrsla, y = lrrtb)) +
  ggtitle("Delta(SLA) vs Log-response ratio of a species' biomass")

# ggplot(visData, aes(x = lsla.d, y = lrrtb)) + 
#   geom_point(aes(size = tot_bio, alpha = 0.05), pch = 19)+
#   geom_point(aes(x = lsla.d, y = predvals, col = "red")) + 
#   geom_smooth(aes(x = lsla.d, y = predvals), method = "lm")

ggpubr::ggarrange(p1,p2)

In general there is some evidence for the change in total biomass to be driven by the sla. Water content and log-response ratios of SLA are significantly affecting the change in total biomass. The change however might be determined by a strong correlation at one site only: Wanang. Higher water content results in positive change in biomass of a species, while if we have large increase in SLA of a species that was correlated with reduction of a species' biomass. **None of the interactions with site were significant.**

The model above has the interactive effect of site, but later I perform analyses at each site separately. **Which approach should I choose?**

However, when we analyse the data separately for each elevation we are getting a different results **if we only use the lsla.d variable**. The more realistic models could change this relationship, i.e., when we include also the effects of species richness and other traits as we did above.

```{r, include=FALSE}
# The Focused model
# Only the predictive power of SLA, water content and percentage herbivory on deltaBio

@@

# dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d,
#                      data = wdata,
#                      method = "ML",
#                      correlation = corMartins(2.718282, 
#                                               phy = myTree.rand.poly, 
#                                               form = ~unified_names,
#                                               fixed = T),
#                      na.action = na.omit)
dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
                     data = wdata,
                     method = "ML",
                     correlation = corMartins(2.718282, 
                                              phy = myTree.rand.poly, 
                                              form = ~unified_names,
                                              fixed = T),
                     na.action = na.omit)

@@

# dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d,
#                      data = ndata,
#                      method = "ML",
#                      correlation = corMartins(2.718282, 
#                                               phy = myTree.rand.poly, 
#                                               form = ~unified_names,
#                                               fixed = T),
#                      na.action = na.omit)
dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
                     data = ndata,
                     method = "ML",
                     correlation = corMartins(2.718282, 
                                              phy = myTree.rand.poly, 
                                              form = ~unified_names,
                                              fixed = T),
                     na.action = na.omit)

@@

# The model
# dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d,
#                      data = ydata,
#                      method = "ML",
#                      correlation = corMartins(2.718282, 
#                                               phy = myTree.rand.poly, 
#                                               form = ~unified_names,
#                                               fixed = T),
#                      na.action = na.omit)

dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
                     data = ydata,
                     method = "ML",
                     correlation = corMartins(2.718282,
                                              phy = myTree.rand.poly,
                                              form = ~unified_names,
                                              fixed = T),
                     na.action = na.omit)

@@

# print the MSE
mse_focused_df <- rbind(mse_focused_df,  
                        data.frame(mse = with(ydata, mean((lrrtb - preds))^2),
                                   treatment = "Insecticide", site = "Yawan"))

library(sjPlot)

@@

# fig.cap = "Empirical and predicted values of a simple model of biomass change vs logarithm of SLA for each studied sites."}

@@

ggpubr::ggarrange(wpredPlot, npredPlot, ypredPlot,
                  nrow = 1)
# Plot the relationship and add predicted values
ggplot(rbind(ydata, ndata, wdata), aes(y = lrrtb, x = lsla.d, color = site))+
  geom_point() + 
  geom_line(data = rbind(ydata, ndata, wdata), 
            aes(x = lsla.d, y = preds), lwd = 1)+
  geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15)

@@

# Fungicide
